\documentclass[10pt]{report}
\usepackage{workbook132}
\input{macros.tex}

\title{Problems and Exercises}
\author{\texttt{CAS CS 132}: \textit{Geometric Algorithms}}
\begin{document}
\maketitle
\tableofcontents

\chapter*{Preface}

The following is a collection of linear algebraic problems and
exercises for the course \texttt{CAS CS 132}: \textit{Geometric
  Algorithms} at Boston University.  As a student of this course, you
will hopefully work through nearly every problem in this document.
The problems are organized by topic and difficulty.  With regard to
difficulty there are four kinds of problems:
\begin{itemize}
\item Basic Exercises
\item True/False
\item More Difficult Problems
\item Challenge Problems
\end{itemize}
You should think of basic exercises as problems that are most like
problems that appear on quizzes, and the true/false and difficult
problems as those that are most like problems that appear on exams
(several of these problems and exercises have appeared on quizzes and
exams in the past).  The challenge problems are for those of who want
to think a bit harder, and are most like problems that may appear as
extra credit on exams.

Problems that require a computer or a calculator are labelled with the
symbol \faCalculator.  Otherwise, the expectation is that you should
be able to solve the problem by hand.

Solutions to these problems will never be made public, except during a
given semester when a problem is given in an assignment.  Also, during
the semester a student of the course is welcome to request a partial
solution any problem via our course public forum (usually Piazza).

Even if it's not explicitly stated, \textbf{you should always show
  your work and justify your answer for every problem.}  Happy
exercising/problem solving!

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Linear Systems}

Determine the coefficient matrix and augmented matrix for each of the
following linear systems.
\begin{enumerate}
\item
  \begin{align*}
    2 x_{1} - 6 x_{2} &= -4 \\
    6 x_{1} + 8 x_{2} &= -7
  \end{align*}
\item
  \begin{align*}
    x_{1} + 8 x_{2} + 7 x_{3} &= -1 \\
    4 x_{1} - 9 x_{2} &= 8 \\
    7 x_{1} - 7 x_{2} - 3 x_{3} &= 10
  \end{align*}
\item
  \begin{align*}
    x_{1} - 2x_{2} - 2x_{3} &= 2 \\
    2x_{1} - 3x_{2} - 5x_{3} &= 2 \\
    -2x_{1} + 2x_{2} + 7x_{3} &= -1
  \end{align*}
\item
  \begin{align*}
    - 8 x_{1} + 6 x_{2} + x_{3} + 7 x_{4} + 10 x_{5} &= 8 \\
    - 4 x_{2} - 4 x_{3} - x_{4} + x_{5} &= -8 \\
    7 x_{1} + x_{2} - 2 x_{3} - 8 x_{4} + 7 x_{5} &= 9 \\
    - 4 x_{1} - 7 x_{2} + 5 x_{3} + 2 x_{4} - 9 x_{5} &= -4
  \end{align*}
\end{enumerate}
For each of the following linear systems, verify that the given point
$s$ is a solution.
\begin{enumerate}[resume]
\item
  \begin{align*}
    x_{1} - 2x_{2} + x_{3} - 2x_{4} &= -9 \\
    x_{1} - x_{2} - x_{3} - 2x_{4} &= -10 \\
    -3x_{1} + 8x_{2} - 6x_{3} + 4x_{4} &= 21 \\
    2x_{2} - 7x_{3} + 7x_{4} &= 13 \\
    \\
    s = (1, 3, 2, 3)
  \end{align*}
\end{enumerate}
Demonstrate that each of the following linear systems has a unique
solution.  Additionally, determine the solution.
\begin{enumerate}[resume]
\item
  \begin{align*}
    3 x_{1} + 6 x_{2} &= -15 \\
    - x_{1} + x_{2} &= 5
  \end{align*}
\item
  \begin{align*}
    2x + 4y &= 29 \\
    -6x - 10y &= -75
  \end{align*}
\item
  \begin{align*}
    x + 3y - 3z &= -2 \\
    2x + 7y + 7z &= -5 \\
    x - 2y - z &= 3
  \end{align*}
\item
  \begin{align*}
    3x - 4y + 3z &= -9 \\
    6x + 7y - 3z &= 0 \\
    x + 10z &= -21
\end{align*}
\item
  \begin{align*}
    - 3 x_{1} + 6 x_{3} &= 12 \\
    3 x_{1} + 3 x_{2} - 12 x_{3} &= -21 \\
    2 x_{1} - 6 x_{2} + 10 x_{3} &= 16
  \end{align*}
\item
  \begin{align*}
    - 3 x_{1} + 6 x_{2} &= 9 \\
    3 x_{1} - 8 x_{2} &= -9 \\
    2 x_{2} + 6 x_{3} &= 18
  \end{align*}
\item
  \begin{align*}
    x_{1} - 2x_{2} - 2x_{3} &= -7 \\
    -x_{1} + 3x_{2} + 2x_{3} &= 10 \\
    2x_{1} - 6x_{2} - 3x_{3} &= -18
  \end{align*}
\item
  \begin{align*}
    - 3 x_{1} - 6 x_{3} + 9 x_{4} &= -6 \\
    - x_{1} + 2 x_{2} - 3 x_{4} &= 8 \\
    - 2 x_{1} - 9 x_{2} - 19 x_{3} + 45 x_{4} &= -61 \\
    x_{1} - 3 x_{2} + 5 x_{3} - 9 x_{4} &= 5
  \end{align*}
\end{enumerate}
For each of the following matrices, apply the given row operations
from top to bottom.
\begin{enumerate}[resume]
\item
  \begin{displaymath}
    \begin{bmatrix}
      -2 & 7 & 9 \\
      7 & -9 & -9 \\
      4 & -1 & 9
    \end{bmatrix}
  \end{displaymath}
  \begin{align*}
    R_{1} &\gets -5R_{1} \\
    R_{1} &\gets R_{1} - 4R_{2} \\
    R_{2} &\gets R_{2} + 3R_{1}
  \end{align*}
\item
  \begin{displaymath}
    \begin{bmatrix}
      -4 & 1 & 3 \\
      -7 & 0 & 5 \\
      -4 & 3 & 1
    \end{bmatrix}
  \end{displaymath}
  \begin{align*}
    R_{3} &\gets R_{3} - 3R_{2} \\
    R_{3} &\gets R_{3} + 3R_{1} \\
    R_{1} &\leftrightarrow R_{3}
  \end{align*}
\item
  \begin{displaymath}
    \begin{bmatrix}
      9 & 5 & -7 & -5 & -9 \\
      5 & -7 & 1 & -2 & -9 \\
      5 & 1 & -10 & 6 & -5 \\
      5 & 7 & -5 & 2 & 1
    \end{bmatrix}
  \end{displaymath}
  \begin{align*}
    R_{4} &\gets -R_{4} \\
    R_{2} &\gets R_{2} - 2R_{3} \\
    R_{2} &\gets R_{2} - 5R_{4} \\
    R_{3} &\gets R_{3} + 3R_{4} \\
    R_{3} &\leftrightarrow R_{2}
  \end{align*}
\end{enumerate}
Determine a general form solution from each of the following
matrices. That is, determine a general form solution for a linear
system whose augmented matrix is row equivalent to the given matrix.
\begin{enumerate}[resume]
\item
  \begin{displaymath}
    \begin{bmatrix}
      1 & -3 & 0 & 0 & -5 \\
      0 & 0 & 1 & 0 & 8 \\
      0 & 0 & 0 & 1 & 5
    \end{bmatrix}
  \end{displaymath}
\item
  \begin{displaymath}
    \begin{bmatrix}
      1 & 0 & -3 & 0 & 1 \\
      0 & 1 & 1 & 1 & 1 \\
      0 & 0 & 0 & 1 & 5
    \end{bmatrix}
  \end{displaymath}
\item
  \begin{displaymath}
    \begin{bmatrix}
      1 & 0 & -6 & 0 & 3 \\
      0 & 1 & -3 & 2 & -9 \\
      0 & 0 & 0 & 1 & -6
    \end{bmatrix}
  \end{displaymath}
\item
  \begin{displaymath}
    \begin{bmatrix}
      1 & 1 & 0 & 0 & 0 & -4 & 5 & -3 \\
      0 & 0 & 1 & 0 & 0 & 1 & 3 & -4 \\
      0 & 0 & 0 & 0 & 1 & 5 & -3 & 2 \\
      0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
      0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
    \end{bmatrix}
  \end{displaymath}
\end{enumerate}
Determine three particular solutions of the linear system underlying
each of the following matrices.  That is, determine three particular
solutions of a linear system whose augmented matrix is row equivalent
to the given matrix.
\begin{enumerate}[resume]
\item
  \begin{displaymath}
    \begin{bmatrix}
      1 & 1 & 0 & -3 & 0 & 0 & -3 & 1 \\
      0 & 0 & 1 & -1 & 0 & 3 & 2 & 3 \\
      0 & 0 & 0 & 0 & 1 & 3 & -4 & -3 \\
      0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
      0 & 0 & 0 & 0 & 0 & 0 & 0 & 0
    \end{bmatrix}
  \end{displaymath}
\end{enumerate}
Determine the row-reduced echelon form of each of the following
matrices.  You must include the intermediate matrices and row
operations used.
\begin{enumerate}[resume]
\item
  \begin{displaymath}
    \begin{bmatrix}
      5 & 2 & 9 \\
      7 & 3 & 12 \\
      2 & 1 & 3
    \end{bmatrix}
  \end{displaymath}
\item
  \begin{displaymath}
    \begin{bmatrix}
      0 & 1 & -6 & -2 \\
      1 & 1 & -3 & -6 \\
      -2 & -4 & 18 & 16
    \end{bmatrix}
  \end{displaymath}
\item
  \begin{displaymath}
    \begin{bmatrix}
      2 & 3 & -13 & -8 \\
      1 & 1 & -4 & -1 \\
      1 & 0 & 1 & 5
    \end{bmatrix}
  \end{displaymath}
\item
  \begin{displaymath}
    \begin{bmatrix}
      1 & -1 & -2 & 1 \\
      -1 & 2 & 4 & 0 \\
      2 & -3 & -6 & 2 \\
      -2 & 1 & 2 & -1
    \end{bmatrix}
  \end{displaymath}
\end{enumerate}
For each of the following linear systems, determine
\begin{itemize}
\item its augmented matrix;
\item the row-reduced echelon form of its augmented matrix;
\item whether it has \textbf{no solutions}, a \textbf{unique
  solution}, or \textbf{infinitely many solutions}.
\end{itemize}
\begin{enumerate}[resume]
\item \faCalculator
  \begin{align*}
    x_1 - 2x_2 + 5x_3 &= 6 \\
    -2x_1 + 6x_2 - 11x_3 + 7x_4 &= -8 \\
    5x_1 - 10 x_2 + 25x_3 + 3x_4 &= 30
  \end{align*}
\item \faCalculator
  \begin{align*}
    x_1 - 3x_2 + 4x_3 &= 0 \\
    -x_1 + 6x_2 + x_3 &= 4 \\
    23x_2 + 5x_3 &= 9
  \end{align*}
\item \faCalculator
  \begin{align*}
    x_1 + 5x_2 + 3x_3 &= -4 \\
    x_1 + 6x_2 + 7x_3 &= -13 \\
    -2x_1 - 12 x_2 - 14 x_3 &= 25
  \end{align*}
\end{enumerate}
Determine a sequence of elementary row operations from the left matrix
to the right matrix.  You must including the intermediate matrices and
row operations used.
\begin{enumerate}[resume]
\item
  \begin{displaymath}
    \begin{bmatrix}
      -4 & -7 & 4 \\
      3 & 8 & 2 \\
      -10 & 1 & -9
    \end{bmatrix}
    \sim
    \begin{bmatrix}
      3 & 8 & 2 \\
      -14 & -6 & -5 \\
      -10 & 1 & -9
    \end{bmatrix}
  \end{displaymath}
\item
  \begin{displaymath}
    \begin{bmatrix}
      1 & 2 & -1 & 5 \\
      0 & 1 & 0 & 1 \\
      0 & 0 & 0 & 5 \\
      2 & 2 & 1 & 1
    \end{bmatrix}
    \sim
    \begin{bmatrix}
      3 & 4 & 0 & 6 \\
      4 & 4 & 2 & 2 \\
      0 & -1 & 0 & 4 \\
      0 & 1 & 0 & 1
    \end{bmatrix}
  \end{displaymath}
\end{enumerate}
Determine a general form solution for each of the following linear
systems by first determining the row-reduced echelon form of its
augmented matrix.  If this system has no solutions then write
\enquote{no solution}.
\begin{enumerate}[resume]
\item
  \begin{align*}
    -x - 2y - z &= -4 \\
    x + 2y &= 3 \\
    -2x -4y + z &= -7
  \end{align*}
\item
  \begin{align*}
    x_{1} - 6x_{2} + x_{4} &= 2 \\
    2x_{1} - 12x_{2} + x_{3} - 4x_{4} &= 7 \\
    x_{1} - 6x_{2} - 2x_{3} + 13x_{4} &= -4
  \end{align*}
\item \faCalculator
  \begin{align*}
    88x_1 + 61x_2 + 8x_3 + 61 x_4 &= -68 \\
    37x_1 + 14x_2 + 72x_3 + 41 x_4 &= -460 \\
    57x_1 + 59x_2 + 69x_3 + 75x_4 &= -333 \\
    92 x_1 + 60x_2 + 28x_3 + 72x_4 &= -192
  \end{align*}
\item \faCalculator
  \begin{align*}
    x_2 + x_3 + x_4 + 3x_5 - x_6 + 2x_7 &= 34 \\
    x_3 - x_4 + 4x_5 &= -14 \\
    2x_2 - 4x_3 + 8x_4 - 18x_5 - x_6 &= 110 \\
    3x_3 - 3x_4 + 12 x_5 + 2x_7 &= -18 \\
    x_2 + 4x_3 - 2x_4 + 15x_5 - x_6 + 7x_7 &= 52
  \end{align*}
\item \faCalculator
  \begin{align*}
    8x_1 + 3x_2 - 3x_3 &= -4 \\
    -3x_1 + 8x_2 + 5x_3 &= -4 \\
    6x_1 + 7x_2 - 3x_3 &= -3 \\
    8x_1 + x_2 - x_3 &= -4
  \end{align*}
\end{enumerate}

\section*{True/False}

\trueFalseHeader

\begin{enumerate}[resume]
\item
  Elementary row operations cannot change the solution set of a linear
  system.
\item
  There is a linear system with exactly three solutions.
\item
  If $A$ is the augmented matrix of an inconsistent linear system, and
  $B$ is a matrix such that $A \sim B$ (that is, $A$ and $B$ are row
  equivalent), then $B$ is the augmented matrix of an inconsistent
  linear system.
\item
  If $A \sim B$ and $A \sim C$ and $B$ and $C$ are in reduced echelon
  form, then $B = C$.
\item
  There is a unique sequence of row operations that reduces a given
  matrix to reduced echelon form.
\item
  If a general form solution of a linear system has a free variable,
  then the system must have infinitely many solutions.
\item
  A matrix may have different pivot positions depending on the
  sequence of row operations used to attain a matrix in echelon form.
\item
  A linear system over 3 variables and 2 equations must be consistent.
\item
  If the coefficient matrix of a linear system as more rows than
  columns, then the system must have infinitely many solutions.
\item
  A \textit{consistent} linear system whose augmented matrix is a $207
  \times 209$ matrix must have infinitely many solutions.
\item
  A \textit{consistent} linear system whose coefficient matrix is
  square must have infinitely many solutions.
\item
  If the rightmost column of the augmented matrix of a linear system
  is a pivot column, then the system is inconsistent.
\end{enumerate}

\section*{More Difficult Problems}
\begin{enumerate}[resume]
\item
  Find a solution to the linear system below with the following
  properties:
  \begin{enumerate}
  \item
    the solution consists entirely of integer values;
  \item
    the values in the solution are relatively prime, i.e., it's not
    possible to divide the solution by a number to get another integer
    solution. So $(2, 4, 6)$ does not satisfy this property but $(1,
    2, 3)$ does.
  \end{enumerate}
  In addition to the solution, you must write down the RREF of the
  augmented matrix of this system, but otherwise, you don't have to
  show your work.\footnote{This is the process we would use to find a
  valid solution to the problem of balancing chemical equations.}
  \begin{align*}
    4x_2 + x_3 &= 16 \\
    9x_1 - 20x_2 - 8x_3 &= -71 \\
    3x_1 - 8x_2 - 3x_3 &= -29
  \end{align*}
\item
  Determine all the $2 \times 2$ matrices in echelon form whose
  entries are either $0$ or $1$.  Mark which ones are in reduced
  echelon form.
\item
  Determine every $3 \times 3$ matrix in reduced echelon form with at
  least two pivot positions whose entries are either $0$ or $1$.
\item
  Determine a linear system over three variables such that $(a, b, c)$
  is a solution exactly when the cubic function $f(x) = ax^3 + bx^2 +
  c$ intersects the points $(-1, 4)$, $(1, 5)$ and $(2, 10)$.  You do
  not need to solve the linear system.
\item
  Determine the slope-intercept form of the line equation which
  defines the intersection of the plane
  \begin{align*}
    2x + 3y + 3z = 6
  \end{align*}
  with the $xy$-plane.
\item
  For what values of the coefficient $h$ is the following system
  inconsistent?
  \begin{align*}
    x + 4y &= -1 \\
    3x - hy &= 7
  \end{align*}
  Is there a value of $h$ for which the above system has infinitely
  many solutions? Justify your answer.
\item
  Consider the following linear system with two unknown coefficients
  $h$ and $k$.
  \begin{align*}
    hx + 2y &= 1 \\
    3x + 9y &= k
  \end{align*}
  \begin{enumerate}
  \item
    Determine values of $h$ and $k$ so that the above linear system
    has no solutions.
  \item
    Determine values of $h$ and $k$ so that the above linear system
    has exactly one solution.
  \item
    Determine values of $h$ and $k$ so that the above linear system
    has infinitely many solutions.
  \end{enumerate}
\item
  Consider the following general form solution
  \begin{align*}
    x_1 &= 3x_2 - 2x_3 \\
    x_2 &\quad \text{is free} \\
    x_3 &\quad \text{is free} \\
    x_4 &= 2 + x_3
  \end{align*}
  \begin{enumerate}
  \item
    Determine another general form solution which describes the same
    solution set, but for which $x_3$ is basic and $x_4$ is free.
  \item
    Determine an RREF which yields the general form solution from the
    previous part.
  \end{enumerate}
\item
  Consider the following general form solution.
  \begin{align*}
    x_{1} &= -6 + 6x_{3} + 2x_{5} \\
    x_{2} &= 4 + 4x_{3} + 6x_{5} \\
    x_{3} &\text{ is free} \\
    x_{4} &= -4 + 5x_{5} \\
    x_{5} &\text{ is free}
  \end{align*}
  Determine a general form solution that describes the same solution
  set but in which $x_1$ is free.
\item
  \faCalculator \ Suppose you're investigating a claim that, out of
  five competing car companies who purchase engine parts from the same
  suppliers, one purportedly \textit{overestimated} their total
  spending by \textdollar 1,300,000.  Companies are required to report
  their total spending, and you've been able to determine how many
  units (× 10,000) of each part that each company has purchased.  You
  haven't been unable to determine how much each item costs per unit
  (it's an industry secret), but you can assume that each company pays
  the same amount per unit.  Given the following data, which company
  is falsifying their records?  Justify our answer.  The total amount
  spent by each company in the table below is multiplied by
  \textdollar 100,000.
  \begin{center}
    \begin{tabular}{|c|r|r|r|r|r|r|}
      \hline
      Co. & Part 1 & Part 2 & Part 3 & Part 4 & Part 5 & Total Spent\\
      \hline
      A & 15 & 3 & 24 & 46 & 182 & 1013 \\
      B & 5 & 3 & 14 & 25 &100 & 552 \\
      C & 15 & 3 & 24 & 46 &188 & 1038 \\
      D & 5 & 0 &  5 & 10 & 40 & 225 \\
      E & 15 & 3 & 26 & 47 &190 & 1056 \\
      \hline
    \end{tabular}
  \end{center}
\item
  Suppose you're given a system of linear equations with three
  variables and two equations, and that $(4, 1, 0)$ and $(2, 4, 1)$
  are solutions to this system.  You are also given that the two
  equations define distinct planes in $\mathbb R^3$.  Determine the
  RREF of the augmented matrix for this system.
\item
  Consider an arbitrary system of linear equations with $n$ unknowns
  and $m$ equations.  Further suppose that
  \begin{itemize}
  \item
    it has a unique solution;
  \item
    it has at least as many equations as unknowns ($m \geq n$).
  \end{itemize}
  Write down an expression in terms of $m$ and $n$ for the number of
  all-zero rows which appear in the rwo-reduced echelon form of its
  augmented matrix.  Justify your answer.
\end{enumerate}

\section*{Challenge Problems}

\begin{enumerate}[resume]
\item
  Consider the following pair of linear systems.
  \begin{align*}
    ax + by &= c \\
    dx + ey &= f
  \end{align*}
  \begin{align*}
    ax + by - cz &= 0 \\
    dx + ey - fz &= 0
  \end{align*}
  \begin{enumerate}
  \item
    Demonstrate that if the first system has a solution, then so does
    the second one.
  \item
    Give \textbf{nonzero} values to $a$ through $f$ such that the
    second system has a solution, but the first does not.  Present
    your solution as an augmented matrix for the first system, i.e.,
    of the form
    \begin{displaymath}
      \begin{bmatrix}
        a & b & c \\
        d & e & f
      \end{bmatrix}
    \end{displaymath}
  \end{enumerate}
\item
  Consider the following pair of linear systems.
  \begin{align*}
    x_1 - 2x_2 &= 3 \\
    4x_1 + x_2 &= 21
  \end{align*}
  \begin{align*}
    10x_3 + 2x_4 &= x_1 \\
    -8x_3 + 9x_4 &= x_2
  \end{align*}
  \begin{enumerate}
  \item
    Solve the first system of linear equations (in $x_1$ and $x_2$)
    and write down the augmented matrix of the second system with the
    solutions of $x_1$ and $x_2$ substituted in.
  \item
    Determine the augmented matrix of a \textit{single} system of
    linear equations \textit{with all four equations} in the variables
    $x_1$, $x_2$, $x_3$, $x_4$.  Describe the relationship between
    this matrix and the one in the previous part.
\end{enumerate}
\item
  Determine what must hold of $a$, $b$, $c$, $d$ so that the following
  linear system has exactly one solution.
  \begin{align*}
    ax + by &= 0 \\
    cx + dy &= 0
  \end{align*}
\item
  Determine an inconsistent linear system in 3 variables such that
  every pair of equations is consistent.  That is, give values for $a$
  through $l$ in the system
  \begin{align*}
    ax + by + cz &= d \\
    ex + fy + gz &= h \\
    ix + jy + kz &= l
  \end{align*}
  such that the system is inconsistent, but each pair of equations
  forms a consistent system.  Present your solution to each part as
  the an augmented matrix, i.e., of the form
  \begin{displaymath}
    \begin{bmatrix}
      a & b & c & d \\
      e & f & g & h \\
      i & j & k & l
    \end{bmatrix}
  \end{displaymath}
  \begin{enumerate}
  \item
    Achieve this with no more than 5 nonzero values for $a$ through
    $l$.
  \item
    Achieve this with all nonzero values for $a$ through $l$.
  \end{enumerate}
\item
  Consider an arbitrary linear system in three variables:
  \begin{align*}
    ax + by + cz &= d \\
    ex + fy + gz &= h \\
    ix + jy + kz &= l
  \end{align*}
  Show that if $(s_x, s_y, s_z)$ and $(t_x, t_y, t_z)$ are solutions
  to the above system, then
  \begin{displaymath}
    \left(
    \frac{s_x + t_x}{2},
    \frac{s_y + t_y}{2},
    \frac{s_z + t_z}{2}
    \right)
  \end{displaymath}
  is also a solution.  Describe why $(s_x, s_y, s_z) \not = (t_x, t_y,
  t_z)$ implies the system has infinitely many solutions.
\item
  One way to describe a line in 3 dimensions is to use a parameter
  $t$.  That is, after fixing values $a_x$, $b_x$, $a_y$, $b_y$,
  $a_z$, and $b_z$, a line can be described as all points of the form
  \begin{displaymath}
    (a_xt + b_x, a_yt + b_y, a_zt + b_z)
  \end{displaymath}
  for any real value of $t$.  Give a pair of 3 dimensional linear
  equations (each of which represents a plane in $\R^3$) whose
  intersection is exactly the line whose points are defined by the
  above parametric form.  \textit{Hint:} The line above can be thought
  of as a system in the variables $x$, $y$, $z$, and $t$, e.g., one of
  its equations is $x - a_xt = b_x$. Write $t$ in terms of $x$ and
  substitute this value for $t$ into the other equations.
\item
  There are several ways to define row equivalence.  For example,
  suppose we restrict the replacement rule to only allow operations of
  the form
  \begin{displaymath}
    R_i \gets R_i + R_j
  \end{displaymath}
  That is, we can only add one row to another, without doing any
  scaling of that row.  We call this an \textit{addition operation}.

  Demonstrate that two matrices $A$ and $B$ are row equivalent if
  there is a sequence of addition and scaling operations which
  transform $A$ to $B$.  In particular, addition and scaling
  operations can simulate replacement and exchange operations.
\end{enumerate}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Vector Equations and Spans}

\section*{Basic Exercises}

Compute the following linear combinations of vectors.
\begin{enumerate}
\item
  \begin{displaymath}
    7 \vFour 2 {-3} {-8} 9
    + 3 \vFour 2 {-3} {-4} {-3}
    - 4 \vFour 5 {-1} 5 {-7}
    - 2 \vFour {-2} {-9} {-2} {-10}
  \end{displaymath}
\end{enumerate}
For each of the following linear systems, determine an equivalent
vector equation, i.e., one that has the same solution set as the given
linear system.
\begin{enumerate}[resume]
\item
  \begin{align*}
    8x_{1} + 6x_{2} - 9x_{3} &= -5 \\
    4x_{1} + 3x_{2} + 9x_{3} + 2x_{4} &= -1 \\
    4x_{1} + 5x_{2} &= 9 \\
    3x_{2} + 8x_{3} - 3x_{4} &= 2
  \end{align*}
\end{enumerate}
Determine a vector that is in $\vspan\{\vv_1, \vv_2\}$ but not in
$\vspan\{\vv_1\}$ or $\vspan\{\vv_2\}$, if possible.  Justify your
answer.  In particular, if it is not possible, then explain why.
\begin{enumerate}[resume]
\item
  \begin{displaymath}
    \vv_1 = \vThree {-7} {-1} 0
    \quad
    \vv_2 = \vThree {-8} 9 {-7}
  \end{displaymath}
\end{enumerate}
Determine a general form solution for each vector equation.  If the
given equation has no solutions then write \enquote{no solution}.
\begin{enumerate}[resume]
\item
  \begin{displaymath}
    x_1 \vThree 1 {-2} 0
    + x_2 \vThree {-2} 5 2
    + x_3 \vThree {-8} {21} {10}
    = \vThree {16} {-38} {-12}
  \end{displaymath}
\end{enumerate}
For each of the following collections of vectors, determine if the
vector $\vv_1$ is in the span of the remaining vectors. If it
is, determine the corresponding dependence relation, i.e., write
$\vv_1$ as a linear combination of the remaining vectors.
\begin{enumerate}[resume]
\item
  \begin{displaymath}
    \vv_1 = \begin{bmatrix} 0 \\ -3 \\ -20 \end{bmatrix}
    \qquad
    \vv_2 = \begin{bmatrix} 1 \\ 1 \\ 3 \end{bmatrix}
    \qquad
    \vv_3 = \begin{bmatrix} 1 \\ 2 \\ 8 \end{bmatrix}
  \end{displaymath}
\item
  \begin{displaymath}
    \vv_1 = \begin{bmatrix} 1 \\ 4 \\ -7 \end{bmatrix}
    \quad
    \vv_2 = \begin{bmatrix} 0 \\ 1 \\ -3 \end{bmatrix}
    \quad
    \vv_3 = \begin{bmatrix} 5 \\ 7 \\ -1 \end{bmatrix}
    \quad
    \vv_4 = \begin{bmatrix} 1 \\ 2 \\ -2 \end{bmatrix}
  \end{displaymath}
\item \faCalculator
  \begin{align*}
    \mathbf v_1
    = \begin{bmatrix} 42 \\ -23 \\ 98 \\ 11 \\ -87 \end{bmatrix}
    \qquad
    \mathbf v_2
    = \begin{bmatrix} 10 \\ -33 \\ -5 \\ -30 \\ -3 \end{bmatrix}
    \qquad
    \mathbf v_3
    = \begin{bmatrix} -47 \\ -1 \\ -2 \\ -25 \\ 24 \end{bmatrix}
  \end{align*}
  \begin{align*}
    \mathbf v_4
    = \begin{bmatrix} 31 \\ -34\\ 11 \\ 39 \\ 25 \end{bmatrix}
    \qquad
    \mathbf v_5
    = \begin{bmatrix}-14 \\ 22 \\ 12 \\ 42 \\ 3\end{bmatrix}
  \end{align*}
\item \faCalculator
  \begin{align*}
    \mathbf v_1
    = \begin{bmatrix} 87 \\ -19 \\ -24 \\ -61 \\ -79 \end{bmatrix}
    \qquad
    \mathbf v_2
    = \begin{bmatrix} 10 \\ -33 \\ -5 \\ -30 \\ -3 \end{bmatrix}
    \qquad
    \mathbf v_3
    = \begin{bmatrix} -47 \\ -1 \\ -2 \\ -25 \\ 24 \end{bmatrix}
  \end{align*}
  \begin{align*}
    \mathbf v_4
    = \begin{bmatrix} 31 \\ -34\\ 11 \\ 39 \\ 25 \end{bmatrix}
    \qquad
    \mathbf v_5
    = \begin{bmatrix}-14 \\ 22 \\ 12 \\ 42 \\ 3\end{bmatrix}
  \end{align*}
\end{enumerate}
For each of the following pairs of vectors $\mathbf v_1$ and $\mathbf
v_2$, determine a linear equation whose point set is the
$\mathrm{span}\{\mathbf v_1, \mathbf v_2\}$.  The linear equation you
determine should have relatively prime integer coefficients (i.e., it
should not be possible to divide the equation by an integer value and
get a new equation with integer coefficients).
\begin{enumerate}[resume]
\item
  \begin{align*}
    \mathbf{v}_{1} = \begin{bmatrix} -2 \\ 1 \\ 2 \end{bmatrix}
    \quad
    \mathbf{v}_{2} = \begin{bmatrix} -1 \\ 0 \\ 2 \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \mathbf v_1 = \begin{bmatrix} 1 \\ 0 \\ -2 \end{bmatrix}
    \qquad
    \mathbf v_2 = \begin{bmatrix} 1 \\ 1 \\ 1 \end{bmatrix}
  \end{align*}
\end{enumerate}

\section*{True/False}
\trueFalseHeader
\begin{enumerate}[resume]
\item
  For any two vectors $\mathbf v_1$ and $\mathbf v_2$ in $\mathbb
  R^n$, there is a vector $\mathbf u$ such that $\mathbf v_1 + \mathbf
  v_2 + \mathbf u = \mathbf 0$.
\item
  For any vectors $\mathbf v_1$, $\mathbf v_2$, and $\mathbf v_3$ in
  $\mathbb R^n$, if $\mathbf v_1 \in \mathsf{span}\{\mathbf v_2,
  \mathbf v_3\}$, then $\mathbf v_2 \in \mathsf{span}\{\mathbf v_1,
  \mathbf v_3\}$.
\item
  The span of any two distinct nonzero vectors in $\mathbb R^3$ is a
  plane.
\item
  For any vector $\mathbf v_1$, $\mathbf v_2$, and $\mathbf v_3$ in
  $\mathbb R^n$, $\mathsf{span}\{\mathbf v_1, \mathbf v_2, \mathbf
  v_3\} = \mathsf{span}\{\mathbf v_1 + \mathbf v_3, \mathbf v_2\}$.
\end{enumerate}

\section*{More Difficult Problems}

\begin{enumerate}[resume]
\item
  Consider the vectors
  \begin{displaymath}
    \mathbf v_1 = \begin{bmatrix} 1 \\ 1 \\ 1 \\ 1 \end{bmatrix}
    \qquad
    \mathbf v_2 = \begin{bmatrix} 1 \\ 1 \\ 2 \\ 2 \end{bmatrix}
  \end{displaymath}
  Write vectors $\mathbf v_3$ and $\mathbf v_4$ such that $\mathbf v_3$ is not in $\mathsf{span}\{\mathbf v_1, \mathbf v_2\}$ and $\mathbf v_4$ is not in $\mathsf{span}\{\mathbf v_1, \mathbf v_2, \mathbf v_3\}$.
\end{enumerate}

\section*{Challenge Problems}

\begin{enumerate}[resume]
\item
  Find a linear equation in three variables whose point set is exactly
  \begin{align*}
    \left\{
    \mathbf v +
    \begin{bmatrix} 2 \\ -3 \\ 2 \end{bmatrix}
    \ \text{where} \
    \mathbf v \in
    \mathsf{span}
    \left\{
    \begin{bmatrix} 1 \\ 1 \\ -5 \end{bmatrix},
    \begin{bmatrix} 0 \\ 1 \\ -1 \end{bmatrix}
    \right\}
    \right\}
  \end{align*}
  In other words, every point in the point set of the equation can be
  expressed as the sum of $[2 \ \ (-3) \ \ 2]^T$ and a vector in the
  span of $[1 \ \ 1 \ \ (-5)]^T$ and $[0 \ \ 1 \ \ (-1)]^T$.
\item
  Consider the following linear equation.
  \begin{displaymath}
    x + y + z = 5
  \end{displaymath}
  The plane represented by this equation does not include the origin;
  such a plane is called \textit{affine}.  Determine vectors $\mathbf
  b$, $\mathbf v_1$, and $\mathbf v_2$ such that every point in the
  plane represented by the above equation (that is, every point $(a,
  b, c)$ such that $a + b + c = 5$) can be written as $\mathbf b +
  \mathbf v$ where $\mathbf v$ is in $\mathsf{span} \{\mathbf v_1,
  \mathbf v_2\}$, i.e.,
  \begin{displaymath}
    \{(x, y, z) : x + y + z = 5\} =
    \{\vb + \vv : \vv \in \vspan\{\vv_1, \vv_2\}\}.
  \end{displaymath}
\end{enumerate}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Matrix Equations}

Compute the matrix-vector multiplication $A\mathbf v$ where $A$ and
$\mathbf v$ are given below.  If it is not possible to multiply $A$
with $\mathbf v$, then explain why.
\begin{enumerate}
\item
  \begin{align*}
    A =
    \begin{bmatrix}
      -10 & 6 & 2 & 8 \\ 1 & 3 & 4 & 5 \\ 0 & -2 & 0 & -9
    \end{bmatrix}
    \quad \mathbf v = \begin{bmatrix} 4 \\ -5 \\ 3 \\ 1 \end{bmatrix}
\end{align*}
\item
  \begin{align*}
    A =
    \begin{bmatrix}
      6 & 1 & -8 & -3 \\ 5 & 0 & -9 & -4
    \end{bmatrix}
    \quad \mathbf v = \begin{bmatrix} -6 \\ 2 \end{bmatrix}
\end{align*}
\end{enumerate}
Determine a general form solution for the matrix equation $A\mathbf x
= \mathbf b$ where $A$ and $\mathbf b$ are given below.  If this
equation has no solutions then write \textit{NO SOLUTION}.
\begin{enumerate}[resume]
\item
  \begin{align*}
    A =
    \begin{bmatrix}
      1 & -3 & 1 & -9 \\ 1 & -2 & 0 & -5 \\ -3 & 8 & -1 & 19 \\ -2 & 4
      & 0 & 10
    \end{bmatrix}
    \quad \mathbf b = \begin{bmatrix} 3 \\ 7 \\ -18
      \\ -14 \end{bmatrix}
\end{align*}
\item \faCalculator
  \begin{align*}
    A =
    \begin{bmatrix}
      1 & 1 & 3 & 1 & -7 & 1 \\ -5 & -4 & -14 & -3 & 24 & -8\\ 0 & -2
      & -2 & -3 & 21 & 3 \\ -6 & -3 & -15 & -2 & 11 & -8
    \end{bmatrix}
    \quad \mathbf b =
    \begin{bmatrix}
      7 \\ -16 \\ -20 \\ -27
    \end{bmatrix}
  \end{align*}
\end{enumerate}
For each of the following matrices, Determine if its columns have full
span, i.e., given a matrix in $\mathbb R^{m \times n}$, determine if
the columns span $\mathbb R^m$.
\begin{enumerate}[resume]
\item
  \begin{align*}
    \begin{bmatrix}
      1 & 2 \\ 2 & 5
    \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \begin{bmatrix}
      1 & -5 & 4\\ -1 & 6 & -3\\ -2 & 13 & -7\\
    \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \begin{bmatrix}
      1 & 2 & 0 & 4 \\ -1 & -1 & 2 & 9 \\ -1 & -2 & 1 & 1 \\ 0 & 2 & 6
      & 36
    \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \begin{bmatrix}
      8 & 3 & 8 \\ 8 & -4 & 4 \\ -2 & 1 & -5 \\ -5 & -7 & -10 \\ -8 &
      9 & 1
    \end{bmatrix}
  \end{align*}
\end{enumerate}

\section*{True/False}

\trueFalseHeader

\begin{enumerate}[resume]
\item
  For $A \in \mathbb R^{m \times n}$ where $m > n$, it is not possible
  for the columns of $A$ to span $\mathbb R^m$.
\item
  For $A \in \mathbb R^{m \times n}$, if $A \mathbf x = \mathbf b$ is
  inconsistent for some vector $\mathbf b$, then it is not possible
  for the columns of $A$ to span $\mathbb R^m$.
\item
  For $A \in \mathbb R^{m \times n}$, if $A \mathbf x = \mathbf 0$ has
  infinitely many solutions, then the columns of $A$ span $\mathbb
  R^m$.
\item
  For $A \in \mathbb R^{m \times n}$, if $A \mathbf x = \mathbf 0$ has
  infinitely many solutions, then $\{ \mathbf v : A\mathbf v = \mathbf
  0\} = \mathbb R^n$.
\item
  For matrices $A$ and $B$ in $\mathbb R^{m \times n}$, let $[A \ B]
  \in \mathbb R ^ {m \times 2n}$ be the matrix obtained by
  horizontally stacking $A$ and $B$.  If $A\mathbf x = \mathbf b$ is
  consistent and $B\mathbf x = \mathbf b$ is consistent, then so is
  $[A \ B] \mathbf x = \mathbf b$.
\item
  For matrices $A$ and $B$ in $\mathbb R^{m \times n}$, if $[A
    \ B]\mathbf x = \mathbf b$ is consistent then $A\mathbf x =
  \mathbf b$ is consistent or $B \mathbf x = \mathbf b$ is consistent.
\item
  For $A \in \mathbb R^{m \times n}$, if $A \mathbf x = \mathbf b$ is
  inconsistent for some vector $\mathbf b$, then $A$ does not have a
  pivot position in every column.
\item
  If $\mathbf x \mapsto A \mathbf x$ is one-to-one and onto, then $A$
  is a square matrix.
\item
  If $\mathbf x \mapsto A \mathbf x$ is neither one-to-one or onto,
  and the reduced echelon form of $A$ only has $0$s and $1$s, then one
  of the columns of $A$ is $\mathbf 0$.
\item
  If the columns of $A$ are linearly independent, then the reduced
  echelon form of $A$ has only $0$s and $1$s.
\item
  If $A$ is an $m \times n$ matrix and $m < n$, then the range of
  $\mathbf x \mapsto A\mathbf x$ is $\mathbb R^m$.
\item
  If the columns of $A \in \mathbb R^{m \times n}$ span all of
  $\mathbb R^m$, then the reduced echelon form of $A$ has only $0$s
  and $1$s.
\end{enumerate}

\section*{More Difficult Problems}

\begin{enumerate}[resume]
\item
  \begin{displaymath}
    A
    \begin{bmatrix}
      1 \\ 1 \\ 1 \\ 1
    \end{bmatrix}
    =
    \begin{bmatrix}
      1 \\ 2 \\ 3 \\ 4
    \end{bmatrix}
  \end{displaymath}
  For each of the following shapes, determine a concrete matrix $A$ so
  that the above equation holds, where $\blacksquare$ represents a
  nonzero entry.  { \newcommand{\bs}{\blacksquare}
    \begin{enumerate}
    \item
      \begin{displaymath}
        \begin{bmatrix}
          \bs&0&0&0 \\ 0&\bs&0&0 \\ 0&0&\bs&0 \\ 0&0&0&\bs
        \end{bmatrix}
      \end{displaymath}
    \item
      \begin{displaymath}
        \begin{bmatrix}
          \bs&0&0&0 \\ \bs&\bs&0&0 \\ \bs&\bs&\bs&0 \\ \bs&\bs&\bs&\bs
        \end{bmatrix}
      \end{displaymath}
    \item
      \begin{displaymath}
        \begin{bmatrix}
          \bs&\bs&\bs&\bs \\ \bs&\bs&\bs&0 \\ \bs&\bs&0&0 \\ \bs&0&0&0
          \\
        \end{bmatrix}
      \end{displaymath}
    \end{enumerate}
  }
\item \faCalculator \ Any three distinct points in the plane define a
  \textit{unique} quadratic equation.  Given three points $(x_1,
  y_1)$, $(x_2, y_2)$ and $(x_3, y_3)$, the equation $y = ax^2 + bx +
  c$ that passes through these three points is given by the solution
  to the following matrix equation.
  \begin{align*}
    \begin{bmatrix}
      x_1^2 & x_1 & 1 \\ x_2^2 & x_2 & 1 \\ x_3^2 & x_3 & 1
    \end{bmatrix}
    \begin{bmatrix} a \\ b \\ c \end{bmatrix}
    =
    \begin{bmatrix} y_1 \\ y_2 \\ y_3 \end{bmatrix}
  \end{align*}
  Use it to determine the unique quadratic equation which passes
  through the points $(2, 13)$, $(3, 25)$ and $(-2, 5)$.
\item
  Determine the RREF of the following matrix in terms of $x$, $y$,
  assuming $x \not = 1$.
  \begin{align*}
    \begin{bmatrix}
      x^2 & x & 1 & y \\ 1 & 1 & 1 & 1 \\ \left(\frac{x +
        1}{2}\right)^2 & \frac{x + 1}{2} & 1 & \frac{y + 1}{2}
    \end{bmatrix}
  \end{align*}
  \textit{Hint:} Don't try to row reduce it. Think in terms of
  polynomial interpolation.
\item
  Let $A$ be a matrix with $n$ rows and 6 columns.  Each row of $A$
  contains the \textbf{unweighted} percentage scores (out of 100) of
  one student on 4 homework assignments (columns 1 through 4) a
  midterm exam (column 5) and a final exam (column 6).
  \begin{displaymath}
    \begin{matrix}
      H_1 & H_2 & H_3 & H_4 & M & F
    \end{matrix}
  \end{displaymath}
  \begin{displaymath}
    \begin{bmatrix}
      p_{11} & p_{12} & p_{13} & p_{14} & p_{15} & p_{16} \\ p_{21} &
      p_{22} & p_{23} & p_{24} & p_{25} & p_{26} \\ \vdots & \vdots &
      \vdots & \vdots & \vdots & \vdots \\ p_{n1} & p_{n2} & p_{n3} &
      p_{n4} & p_{n5} & p_{n6} \\
    \end{bmatrix}
  \end{displaymath}
  All homework assignments are worth the same amount.  Let $T$ denote
  the linear transformation implemented by this matrix.
  \begin{enumerate}
  \item
    Suppose that homework assignments account for \textbf{50} percent
    of the final grade, the midterm exam accounts for \textbf{20}
    percent and the final exam accounts for \textbf{30} percent.  Find
    a vector $\mathbf v$ such that $T(\mathbf v)$ is the vector whose
    $i^\text{th}$ entry is the final percentage grade of the
    $i^\text{th}$ student.  For example, if the $i^\text{th}$ student
    recieved $90$ percent on every homework assignment, $85$ percent
    on the midterm, and $92$ percent on the final, then the
    $i^\text{th}$ entry of the output vector should be $90 * 0.5 + 85
    * 0.2 + 92 * 0.3$.
  \item
    Find a vector $\mathbf v$ such that $T(\mathbf v)$ is the vector
    whose $i^\text{th}$ entry is the unweighted homework grade for
    student $i$. For the same example as above, the $i^\text{th}$
    entry would be $90$.
  \end{enumerate}
\end{enumerate}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Linear Independence}

\section*{Basic Exercises}

Determine if the following collections of vectors are linearly
dependent.  If they are write a dependence relation, i.e., determine
linear combination of the given vectors which sums to $\mathbf 0$.
\begin{enumerate}
\item
  \begin{align*}
    \mathbf{v}_{1} = \begin{bmatrix} 1 \\ -1 \end{bmatrix} \quad
    \mathbf{v}_{2} = \begin{bmatrix} -1 \\ 2 \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \mathbf v_1 =
    \begin{bmatrix}
      1 \\ -1 \\ 2
    \end{bmatrix}
    \qquad \mathbf v_2 =
    \begin{bmatrix}
      -1 \\ 4 \\ -3
    \end{bmatrix}
    \qquad \mathbf v_3 =
    \begin{bmatrix}
      -3 \\ 9 \\ -8
    \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \mathbf{v}_{1} = \begin{bmatrix} 1 \\ -2 \\ 1 \end{bmatrix} \quad
    \mathbf{v}_{2} = \begin{bmatrix} -3 \\ 7 \\ -2 \end{bmatrix} \quad
    \mathbf{v}_{3} = \begin{bmatrix} -6 \\ 15 \\ -3 \end{bmatrix}
    \quad \mathbf{v}_{4} = \begin{bmatrix} -1 \\ 3 \\ 1 \end{bmatrix}
  \end{align*}
\end{enumerate}

\section*{True/False}
\trueFalseHeader
\begin{enumerate}[resume]
\item
  For $A \in \mathbb R^{m \times n}$, if the columns of $A$ are
  linearly dependent, then they do not span $\mathbb R^m$.
\end{enumerate}

\section*{More Difficult Problems}

\begin{enumerate}[resume]
\item
  Consider four vectors $\mathbf v_1$, $\mathbf v_2$, $\mathbf v_3$
  and $\mathbf v_4$ in $\mathbb R^4$ with the property that
  \begin{align*}
    \begin{bmatrix}
      \mathbf v_1 & \mathbf v_2 & \mathbf v_3 & \mathbf v_4
    \end{bmatrix}
    \sim
    \begin{bmatrix}
      1 & 1 & 3 & -2 \\ 0 & 4 & -4 & 12 \\ 0 & 0 & 3 & 3 \\ 0 & 0 & 0
      & 0
    \end{bmatrix}
  \end{align*}
  \begin{enumerate}
  \item
    Determine if $\mathbf v_4 \in \mathrm{span}\{\mathbf v_1, \mathbf
    v_2, \mathbf v_3\}$.  If so, express $\mathbf v_4$ as a linear
    combination of the vectors $\mathbf v_1$, $\mathbf v_2$ and
    $\mathbf v_3$.
  \item Deterime if vectors $\mathbf v_2$, $\mathbf v_3$, and $\mathbf
    v_4$ linearly independent.  If so, justify your answer.  If not,
    determine a dependence relation for these vectors.
  \end{enumerate}
\item
  Determine all values of $h$ for which the following set of vectors
  is linearly dependent.
  \begin{align*}
    \mathbf v_1 =
    \begin{bmatrix} 1 \\ 9 \\ -3 \end{bmatrix}
    \qquad \mathbf v_2 =
    \begin{bmatrix} 1 \\ 4 \\ 3 \end{bmatrix}
    \qquad \mathbf v_3 =
    \begin{bmatrix} -2 \\ h \\ -6 \end{bmatrix}
    \qquad
  \end{align*}
\item
  Determine three \textit{nonzero} vectors $\mathbf v_1$, $\mathbf
  v_2$, and $\mathbf v_3$ in $\mathbb R^3$ such that
  \begin{itemize}
  \item
    $\{\mathbf v_1, \mathbf v_2, \mathbf v_3\}$ is linearly dependent
    and
  \item
    $\mathbf v_1$ cannot be written as a linear combination of
    $\mathbf v_2$ and $\mathbf v_3$.
  \end{itemize}
\item \faCalculator \ Consider the following matrix, presented as a
  SymPy array.
  \begin{lstlisting}

  import sympy a = sympy.Matrix( [[ 13, -19, 19, 16, 5, 1, 10, 5, 15],
    [-11, -7, -10, 7, 2, -8, -10, -19, 6], [ -5, 10, -7, 2, -8, 2,
      -15, -16, -11], [ 17, -13, 9, 13, 19, 8, -3, -9, 0], [ 9, -18,
      5, 1, 4, 14, 9, 8, -4], [ 8, 14, 17, 5, -6, 7, -13, 2, 12], [
      18, 12, -7, 2, -10, 15, -12, 1, -12], [ 19, -12, 1, -16, 2, -6,
      -4, 17, 15], [-19, -6, -16, -20, -20, -3, 7, 3, 14], [ 6, 8,
      -15, 5, -5, 8, -14, 5, -19]])
  \end{lstlisting}
  \begin{enumerate}
  \item Determine if the columns of this matrix span all of $\mathbb
    R^{10}$. Justify your answer.
  \item Determine if the columns of this matrix linearly
    independent. Justify your answer.
  \end{enumerate}
\item
  Consider the following collection of vectors.
  \begin{align*}
    &\mathbf v_1 =
    \begin{bmatrix} -10 \\ -6 \\ 2 \\ 0 \\ 8 \\ -3 \\ -4 \end{bmatrix}
    \qquad \mathbf v_2 =
    \begin{bmatrix} -3 \\ 3 \\ -8 \\ 5 \\ -2 \\ 7 \\ -3 \end{bmatrix}
    \qquad \mathbf v_3 =
    \begin{bmatrix} -20 \\ -12 \\ 4 \\ 0 \\ 16 \\ -6 \\ -8 \end{bmatrix}
    \qquad \mathbf v_4 =
    \begin{bmatrix} 6 \\ 1 \\ 3 \\ -1 \\ 10 \\ 7 \\ 4 \end{bmatrix}
  \end{align*}
  \begin{align*}
    &\mathbf v_5 =
    \begin{bmatrix} -2 \\ -10 \\ 10 \\ -5 \\ -3 \\ -10 \\ -5 \end{bmatrix}
    \qquad \mathbf v_6 =
    \begin{bmatrix} -2 \\ 3 \\ -9 \\ -2 \\ -9 \\ -2 \\ 7 \end{bmatrix}
    \qquad \mathbf v_7 =
    \begin{bmatrix} -13 \\ 25 \\ -22 \\ 13 \\ 24 \\ 41 \\ 15 \end{bmatrix}
  \end{align*}
  \begin{enumerate}
  \item
    Determine the first vector (i.e., the vector with the smallest
    index) which can be written as a linear combination as of the
    vectors which precede it.  Express this vectors as a linear
    combination of the vectors which precede it.  You should write
    your final solution using the vector names of the form $\mathbf
    v_i$.
  \item
    Determine a dependence relation for the entire set of vectors with
    the following properties:
    \begin{itemize}
    \item
      the coefficients of the dependence relation are relatively prime
      integers;
    \item
      the number of nonzero coefficients is maximum, i.e., there is no
      other dependence relation with a greater number of nonzero
      coefficients.
    \end{itemize}
  \end{enumerate}
\end{enumerate}

\section*{Challenge Problems}

\begin{enumerate}[resume]
\item
  Determine all values of $h$ for which the following set of vectors
  is linearly dependent.
  \begin{align*}
    \mathbf v_1 =
    \begin{bmatrix} 1 \\ -1 \\ h \end{bmatrix}
    \qquad \mathbf v_2 =
    \begin{bmatrix} 4 \\ 0 \\ 2 \end{bmatrix}
    \qquad \mathbf v_3 =
    \begin{bmatrix} 13 \\ h \\ -1 \end{bmatrix}
    \qquad
  \end{align*}
\end{enumerate}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Linear Transformations}
\input{linear-transformations.tex}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Matrix Algebra}
\input{matrix-algebra.tex}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{LU Factorization}
\input{lu.tex}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Markov Chains}
\input{markov-chains.tex}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Vector Spaces}
\input{vector-spaces.tex}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Eigenvectors}
\input{eigenvectors.tex}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Diagonalization}
\input{diagonalization.tex}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Analytic Geometry and Orthogonality}
\input{analytic-geometry.tex}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Least Squares and Linear Models}
\input{least-squares.tex}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Quadratic Forms}
\input{quadratic-forms.tex}

% ======================================================================
% ======================================================================
% ======================================================================

\chapter{Singular Value Decomposition}
\input{svd.tex}

\end{document}
