\section*{Basic Exercises}

For each of the following stochastic matrices $A$, answer the
following items:
\begin{itemize}
\item
  Draw the state diagram for the corresponding Markov chain. You
  should label each node with a positive integer for the corresponding
  column of $A$ (e.g, the node corresponding to the first column
  should be labeled \enquote{1}, to the second column labeled
  \enquote{2}, and so on).
\item
  Determine if $A$ is regular, and write down the smallest $k$ such
  that $A^k$ has strictly positive values. If it is not regular,
  justify your answer.
\item
  Determine the general form solution of the equation
  $(A-I)\,\mathbf{x} = \mathbf{0}$. You may use a computer to do this,
  but you should express your answer in fractions, not decimals.
\item
  Determine a steady state vector for $A$. If the steady state vector
  is unique, note this. You must do this by hand and show your work.
\end{itemize}
\begin{enumerate}
\item
  \begin{align*}
    \begin{bmatrix}
      0.4 & 0.8 \\ 0.6 & 0.2
    \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \begin{bmatrix}
      0.6 & 0 & 0.1 \\
      0.4 & 0.3 & 0 \\
      0 & 0.7 & 0.9
    \end{bmatrix}
  \end{align*}
\item
  \begin{displaymath}
    \begin{bmatrix}
      1 & 1/2 & 1/3 \\
      0 & 1/2 & 1/3 \\
      0 & 0 & 1/3
    \end{bmatrix}
  \end{displaymath}
\item
  \begin{displaymath}
    \begin{bmatrix}
      0.9 & 0.9 \\
      0.1 & 0.1
    \end{bmatrix}
  \end{displaymath}
\item
  \begin{align*}
    \begin{bmatrix}
      0.2 & 0 & 0 \\
      0.4 & 1 & 0.3 \\
      0.4 & 0 & 0.7
    \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \begin{bmatrix}
      1 & 0 & 0 \\
      0 & 0 & 1 \\
      0 & 1 & 0
    \end{bmatrix}
  \end{align*}
\item
  \begin{displaymath}
    \begin{bmatrix}
      0.2 & 0 & 0.3 \\
      0.8 & 0.5 & 0 \\
      0 & 0.5 & 0.7
    \end{bmatrix}
  \end{displaymath}
\item
  \begin{align*}
    \begin{bmatrix}
      0 & 0 & 0.3 \\
      1 & 0 & 0.7 \\
      0 & 1 & 0
    \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \begin{bmatrix}
      0.9 & 0.6 \\
      0.1 & 0.4
    \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \begin{bmatrix}
      1 & 0.3 & 0.4 \\
      0 & 0.7 & 0.1 \\
      0 & 0 & 0.5
    \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \begin{bmatrix}
      0 & 0.5 & 0.5 \\
      0.5 & 0 & 0.5 \\
      0.5 & 0.5 & 0
    \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \begin{bmatrix}
      0 & 0 & 1 \\
      1 & 0 & 0 \\
      0 & 1 & 0
    \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \begin{bmatrix}
      0 & 0 & 0 & 0.5 \\
      1 & 0 & 0 & 0.3 \\
      0 & 1 & 0 & 0.2 \\
      0 & 0 & 1 & 0
    \end{bmatrix}
  \end{align*}
\item
  \begin{align*}
    \begin{bmatrix}
      0.4 & 0.8 & 0.5 & 0 \\
      0.6 & 0.2 & 0 & 0 \\
      0 & 0 & 0 & 0 \\
      0 & 0 & 0.5 & 1
    \end{bmatrix}
  \end{align*}
\end{enumerate}


\section*{True/False}

\trueFalseHeader

\begin{enumerate}[resume]
\item
  If $A$ is a stochastic matrix, then the all ones vector $\mathbf 1$
  is an eigenvector for $A^T$.
\item
  If a stochastic matrix has a unique steady state vector, then the
  corresponding Markov chain converges to it regardless of starting
  state.
\item
  A stochastic regular matrix $A \in \mathbb R^{n \times n}$ has a
  smallest integer $k$ such that $A^k$ has strictly positive
  entries. For such a matrix, $k \leq n$.
\item
  Every stochastic matrix has at least one steady state vector.
\item
  If $A$ and $B$ are stochastic matrices, then so is $AB$.
\item
  Every $A$ stochastic matrix is invertible.
\end{enumerate}

\section*{More Difficult Problems}

\begin{enumerate}[resume]
\item
  Suppose you've been watching the stock price of your favorite
  company and you've discovered the following trends based on the data
  you've taken:

  \begin{itemize}
  \item
    If the price is \textbf{STEADY} there's a 70\% chance it will
    remain steady the next day, but a 20\% chance it will go
    \textbf{HIGH} and a 10\% chance it will dip \textbf{LOW}.
  \item
    If the price is \textbf{HIGH} there's a 55\% chance it will remain
    high the next day, but a 40\% chance it will go back to being
    \textbf{STEADY}, and just a 5\% chance it will dip all the way
    down to \textbf{LOW}.
  \item
    If the price is \textbf{LOW}, there is a 60\% change it will
    remain low the next day, but a 30\% chance it will reset to
    \textbf{STEADY}, and a 10\% chance it will suddenly jump to
    \textbf{HIGH}.
  \end{itemize}

  Suppose that the price is currently \textbf{STEADY}.  In the long
  term, is more likely to remain \textbf{STEADY}, go up \textbf{HIGH},
  or dip down \textbf{LOW}?  Justify your answer.

\item
  Suppose that we're trying to predict the preformance of the $A_5$
  Soccer Club based on statistics that we've gathered on their recent
  games. We've found that:
  \begin{itemize}
  \item
    After a win, they have a 70\% chance of winning, a 15\% chance of
    drawing, and a 15\% chance losing their next game.
  \item
    After a draw, they have a 20\% chance of winning, a 50\% chance of
    drawing, and a 30\% chance of losing their next game.
  \item
    After a loss, they have a 20\% chance of winning, a 70\% chance of
    drawing, and a 10\% chance of losing their next game.
  \end{itemize}
  \begin{enumerate}
  \item
    If they win their first game, how should we expected them to
    perform overall?  That is, what is the expected percentage of
    wins, draws, and losses in the long term?
  \item
    If they lose their first game, what will their overall record tend
    towards?
  \end{enumerate}
\end{enumerate}

\section*{Challenge Problems}

\begin{enumerate}[resume]
\item
  Demonstrate that any invertible stochastic matrix $A$ with a
  stochastic inverse must be a permutation matrix.  That is, every
  column and every row of $A$ has a single 1, with the rest of the
  entries being 0s.  \textit{Hint:} Consider what can be inferred
  about the entries of a stochastic $A^{-1}$ from its action on $A
  \mathbf{e}_i$ for any elementary basis vector $\mathbf{e}_i$.
\end{enumerate}
