The following is a collection of linear algebraic problems and
exercises for the course \texttt{CAS CS 132}: \textit{Geometric
  Algorithms} at Boston University.  As a student of this course, you
will hopefully work through nearly every problem in this document.
The problems are organized by topic and difficulty.  With regard to
difficulty there are four kinds of problems:
\begin{itemize}
\item Basic Exercises
\item True/False
\item More Difficult Problems
\item Challenge Problems
\end{itemize}
You should think of basic exercises as problems that are most like
problems that appear on quizzes, and the true/false and difficult
problems as those that are most like problems that appear on exams
(several of these problems and exercises have appeared on quizzes and
exams in the past).  The challenge problems are for those of who want
to think a bit harder, and are most like problems that may appear as
extra credit on exams.

Problems that require a computer or a calculator are labelled with the
symbol \faCalculator.  Otherwise, the expectation is that you should
be able to solve the problem by hand.

Solutions to these problems will never be made public, except during a
given semester when a problem is given in an assignment.  Also, during
the semester a student of the course is welcome to request a partial
solution any problem via our course public forum (usually Piazza).

Even if it's not explicitly stated, \textbf{you should always show
  your work and justify your answer for every problem.}  Happy
exercising/problem solving!
